---
title: "Class10 Markdown"
author: "Tom Quach (A15549142)"
date: "10/28/2021"
output: html_document
---

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy <- read.csv(url, row.names = 1)
head(candy, n =5)
```


> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

```{r}
sum(candy$chocolate)
```

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Almond Joy",]$winpercent
```


> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, the winpercent variable seem to be on a different scale 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

I thnk that zero and one are logical variables where zero represents false and 1 represents true 

Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

> Q9. Is the distribution of winpercent values symmetrical?

No

> Q10. Is the center of the distribution above or below 50%?

Below


> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

First need to find all the chocolate candy rows in the "candy" data

```{r}
inds <- as.logical(candy$chocolate)
chocolate <- mean(candy[inds,]$winpercent)

infs <- as.logical(candy$fruity)
fruity <- mean(candy[infs,]$winpercent)

mean(chocolate)
mean(fruity)
```
On average, chocolate is ranked higher than fruity candy


> Q12. Is this difference statistically significant?

```{r}
t.test(candy[inds,]$winpercent, candy[infs,]$winpercent)
```

The difference is statistically significant because p < 0.05

> Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```



>Q14. What are the top 5 all time favorite candy types out of this set?





> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent))  +
  geom_col()
```

Setup a color vetor for coloring candy by whether it is chocolate/fruit/bar..etc

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

> Q17. What is the worst ranked chocolate candy?

The worst ranking chocolate will be sixlets

> Q18. What is the best ranked fruity candy?

Starburst


```{r}
library(ggrepel)
# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reeses miniatures

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

The top 5 most expensive candy types are: Nik L Nip, Nestle Smarties, Ring Pop, HersheyA s Krackle, and Hershey A s Milk Chocolate. The least popular is Nik L Nip

> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

```{r}
library(corrplot)
```
```{r}
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Chocolate and fruity


>Q23. Similarly, what two variables are most positively correlated?

Winpercent and chocolate


**PCA**

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)
```


```{r}
plot(pca$x[,1:2], col=my_cols)
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```


```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
```















