---
title: 'Class 19: Boxplot'
author: "Tom Quach (A15549142)"
date: "12/5/2021"
output:
  pdf_document: default
  html_document: default
---

## Section 4 Population Scale Analysis

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

**Reading our file**

```{r}
expr <- read.table("Expression genotype results.txt")
head(expr)
```

```{r}
nrow(expr)
```

**Determining the sample size for each genotype**

```{r}
table(expr$geno)
```

**Finding the median expression levels for each of these genotypes**

```{r}
summary(expr)
```


```{r}
median(expr$exp)
```
> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
library(ggplot2)
```


##Lets make the boxplot

```{r}
ggplot(expr) + aes(x=geno, y=exp, fill=geno) +geom_boxplot(notch=TRUE)
```

From the boxplot, we can see that the expression value between AA and GG are pretty different, with G/G expression being lower compared to A/A expression. It also looks like having a G/G in this location is definitely associated with the expression of ORMDL3. In sums, the SNP does effect the expression of ORMDL3. 
