---
title: "Machine Learning"
author: "Tom Quach (A15549142)"
date: "10/21/2021"
output:
  pdf_document: default
  html_document: default
---

First up is clustering methods 

#Kmeans clustering

The function in base R to do Kmeans clustering is called 'kmeans()'

First make up some data where we know what the answer should be:

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

> Q. Can we use kmeans() to cluster this data setting k 2 and nstart to 20?

```{r}
km <- kmeans(x, centers = 2, nstart=20)
km
```


> Q. How many points are in each cluster? 

```{r}
km$size
```

> Q. What ‘component’ of your result object details
 - cluster size?

```{r}
km$cluster
```
 
 - cluster assignment/membership?
 - cluster center?
 
```{r}
km$centers
```
 

> Q. Plot x colored by the kmeans cluster assignment and add cluster centers as blue points


```{r}
plot(x, col=km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```



#Hierarchical Clustering

A big limitation with k-means is that we have to tell it K(the number of clusters we want).

Analyze this same data with hclust()

Demonstrate the use of dist(), hclust(), plot(), and cutree(), functions to do clustering. Generate dendrograms and return cluster assignment membership vector

```{r}
hc <- hclust( dist(x) )
hc
```


There is a plot method for hclust result objects. Let's see it 

```{r}
plot(hc)
```

To get our cluster membership vector we have to do a wee bit more work. We have to "cut" the tree where we think it makes sense. For this we use the 'cutree()' function.

```{r}
cutree(hc, h=6)
```

You can also call 'cutree()' setting k=the number of grps/clusters you want. 

```{r}
grps <- cutree(hc, k=2)
```

Make our results plot

```{r}
plot(x, col=grps)
```


##PCA Lab Report

Importing and reading the data

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```


> Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
dim(x)
```
>Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

We can see that there are 17 rows and 5 columns in my new data frame x 

There should only be a 17 x 4 dimensions. We have an extra first column that we need to fix 

```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

As we run the code each time, we are deleting more data from our data frame which is really dangerous 



```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names = 1)
head(x)
```


Rechecking the dimension of the data fram
```{r}
dim(x)
```

**Perfect! The better way seems to be the latter code we just performed, 'row.name()=1'. The first method will delete our data each time we run it**

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

> Q3: Changing what optional argument in the above barplot() function results in the following plot?

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

> Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?

```{r}
pairs(x, col=rainbow(10), pch=16)
```


**Each individual graphs plot two countries against one another in the x and y axis. The y axis show a constant country while the x axis has different countries. If the point lies in the diagonal line, that means there is less variance of that specific point between two countries**

>Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?

**For the other countries, the points are more evenly distributed along an imaginary diagonal line. However, the data set with N. Ireland is more clustered towards the bottom left of each data graph.**

#PCA to the rescue

The main function in base R for PCA is 'prcomp()'

```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
attributes(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2])
```


>Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```


> Q8. Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document.

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500), )
text(pca$x[,1], pca$x[,2], colnames(x), col=c("Orange", "Red", "Blue", "Green"))
```

We are now finding the standard deviation of our data frame

```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```
We are now going to store the important components of pca into a vector
```{r}
z <- summary(pca)
z$importance
```
Creating a plot to see which principal component has the most variance 

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```


Generating a loading plot for pc1 to see what is causing the high vairance
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```


> Q9: Generate a similar ‘loadings plot’ for PC2. What two food groups feature prominantely and what does PC2 maninly tell us about?

Generating a loadings plot for pc2
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```

**The two food groups feature prominately are fresh potatoes and soft drinks. This means that soft drinks, a more negative loading score, are pushing countries towards the bottom side of the plot. This also means that fresh potatoes, a more positive loading score, is pushing countires towards the upper side of the plot**

##PCA of RNA

Inserting and reading RNA-sequence file

```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```

>Q10: How many genes and samples are in this data set?

```{r}
dim(rna.data)
```
**There are a 100 genes and 10 samples**

Generating a barplot

```{r}
## Again we have to take the transpose of our data 
pca <- prcomp(t(rna.data), scale=TRUE)
 
## Simple un polished plot of pc1 and pc2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
```
  
  Getting the important details from the pca of RNA-seq
  
```{r}
summary(pca)
```
  

Creating a plot to see which principal componet has what variance

```{r}
plot(pca, main="Quick scree plot")
```

Looking at the exact value of the variance for each principle component 

```{r}
## Variance captured per PC 
pca.var <- pca$sdev^2

## Percent variance is often more informative to look at 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```

Titling our bar plot 

```{r}
barplot(pca.var.per, main="Scree Plot", 
        names.arg = paste0("PC", 1:10),
        xlab="Principal Component", ylab="Percent Variation")
```

Plotting our PC1 and PC2 against each other to see the majority of the variance more clearly 

```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(rna.data)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"

plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))

text(pca$x[,1], pca$x[,2], labels = colnames(rna.data), pos=c(rep(4,5), rep(2,5)))
```

Using ggplot to create a scatter plot of the rna-sequence data

```{r}
library(ggplot2)

df <- as.data.frame(pca$x)

# Our first basic plot
ggplot(df) + 
  aes(PC1, PC2) + 
  geom_point()
```

Naming each point to see if they are wild type or knock out 

```{r}
# Add a 'wt' and 'ko' "condition" column
df$samples <- colnames(rna.data) 
df$condition <- substr(colnames(rna.data),1,2)

p <- ggplot(df) + 
        aes(PC1, PC2, label=samples, col=condition) + 
        geom_label(show.legend = FALSE)
p
```

Labeling my ggplot

```{r}
p + labs(title="PCA of RNASeq Data",
       subtitle = "PC1 clealy seperates wild-type from knock-out samples",
       x=paste0("PC1 (", pca.var.per[1], "%)"),
       y=paste0("PC2 (", pca.var.per[2], "%)"),
       caption="BIMM143 example data") +
     theme_bw()
```




