---
title: "Class 17 Vaccination Rate Mini Project"
author: "Tom Quach (A15549142)"
date: "11/23/2021"
output:
  pdf_document: default
  html_document: default
---

Download data from CA.gov

```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

> Q1. What column details the total number of people fully vaccinated?

The persons fully vaccinated column details the total number of people fully vaccinated

> Q2. What column details the Zip code tabulation area?

The zip code tabulation area column details the zip code tabulation area


```{r}
head(vax$as_of_date)
```
> Q3. What is the earliest date in this dataset?

2021-01-05

```{r}
tail(vax$as_of_date)
```
> Q4. What is the latest date in this dataset?

2021-11-16

##Skirm


library(skimr)

skimr::skim(vax)

> Q5. How many numeric columns are in this dataset?

There are 9 numeric columns in this dataset

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

There are 8256 "missing values" for person_fully_vaccinated for this dataset

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
8256/81144
```
10.02% are missing

> Q8. [Optional]: Why might this data be missing?

The data is missing may be the result of people are still not fully vaccinated yet


## Ensure the date column is useful

We will use the **lubridate** package to make life a lot easier when dealing with dates and times: 

```{r}
library(lubridate)
```

```{r}
today()
```

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```


```{r}
today() - vax$as_of_date[nrow(vax)]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

> Q9. How many days between the first and last entry in the dataset?

It has been 7 days between the first and last entry in the dataset 

```{r}
length(unique(vax$as_of_date))
```
> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

There are a total of 46 unique dates in the dataset 


##Working with zip codes

```{r}
library(zipcodeR)
```
```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
# Subset to San Diego county only areas
sd <- vax[vax$county == "San Diego",]
```


```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```


## Focus on San Diego findings

> Q11. How many distinct zip codes are listed for San Diego County?


```{r}
length(unique(sd$zip_code_tabulation_area))
```
There are 107 distinct zip codes listed for San Diego County

> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
ind <- which.max(sd$age12_plus_population)
sd[ind,]
```
**The ip code area 92154 in SD has the largest 12+ population in this dataset**

What is the population 

```{r}
filter(sd, zip_code_tabulation_area == "92037")[1,]
```

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

```{r}
sd.now <- filter(sd, as_of_date == "2021-11-09")
head(sd.now)
```
```{r}
mean(sd.now$percent_of_population_fully_vaccinated, na.rm=TRUE)
```
We can look at the 6-number summary

```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```

**The overall average is 0.6727567**

> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

```{r}
library(ggplot2)

ggplot(sd.now) + aes(sd.now$percent_of_population_fully_vaccinated) + geom_histogram(bins=15)
```


What about 92037 - UCSD/La Jolla

##Time series of vaccination rate for 92037

First select all data for the UCSD 92037

```{r}
ucsd <- filter(sd, zip_code_tabulation_area == "92037")
ucsd[1,]$age5_plus_population
```
```{r}
head(ucsd)
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```
> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated") +
  geom_hline(yintercept = 0.66)
```


```{r}
ucsd[1,]$age5_plus_population
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 & as_of_date == "2021-11-16")

summary(vax.36)
```

> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) + aes(vax.36$percent_of_population_fully_vaccinated) + geom_histogram()
```


> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```

92040 is under and 92109 is above

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144

First we need to subset the full "vax" dataset to include only ZIP code areas with a population as large as 92037

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)
nrow(vax.36.all)
```

```{r}
length(unique(vax.36.all$zip_code_tabulation_area))
```

Let's make a final figure that shows all these ZIP areas

```{r}
ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  geom_hline(yintercept = 0.66, col="red") +
  labs(x="Date", y="Percent of Area Vaccinated")
```





























